Reducing iptables configuration complexity using chains

### 2014-05-03 iptables/paper.tex
add configuration

\section{Optimizing configuration with an example setup}
\section{Conclusion}

>>>>

\section{Optimizing configuration with an example setup}
\subsection{Actual configuration}

\subsubsection{INPUT chain}

\begin{verbatim}
  iptables -A INPUT -i lo -j ACCEPT
  iptables -A INPUT -m state --state ESTABLISHED -j ACCEPT
  iptables -A INPUT -m state --state RELATED -j ACCEPT
  iptables -A INPUT -p icmp -j icmp_in
  iptables -A INPUT -p tcp -m tcp --dport 22 -j admin_IP
  iptables -A INPUT -p tcp -m tcp --dport 10000 -j admin_IP
  iptables -A INPUT -p tcp -m tcp --dport 80 -j ACCEPT
  iptables -A INPUT -p tcp -m tcp --dport 443 -j ACCEPT
  iptables -A INPUT -p tcp -m tcp --dport 139 -j webmaster_IP
  iptables -A INPUT -p tcp -m tcp --dport 445 -j webmaster_IP
  iptables -A INPUT -j DROP
\end{verbatim}

This is the INPUT chain, allowing :
\begin{itemize}
  \item all local traffic (not leaving the physical PC)
  \item established and related connections
  \item ICMP packets (ping, etc.) are handled in a seperate chain icmp\_in
  \item some services
  \begin{itemize}
    \item ssh (tcp 22) and webmin (tcp 10000) allowed for admins (admin\_IP chain)
    \item website (tcp 80 and 443) for everybody
    \item samba (tcp 139 and 445) for webmasters (and admins, see definition of webmaster\_IP chain)
  \end{itemize}
  \item everything else is not allowed (dropped)
\end{itemize}

\subsubsection{OUTPUT chain}

\begin{verbatim}
  iptables -A OUTPUT -o lo -j ACCEPT
  iptables -A OUTPUT -m state --state ESTABLISHED -j ACCEPT
  iptables -A OUTPUT -m state --state RELATED -j ACCEPT
  iptables -A OUTPUT -p icmp -m icmp --icmp-type 8 -j ACCEPT
  iptables -A OUTPUT -p udp -m udp --dport 53 -j ACCEPT
  iptables -A OUTPUT -p udp -m udp --dport 67 -j ACCEPT
  iptables -A OUTPUT -p tcp -m tcp --dport 25 -j ACCEPT
  iptables -A OUTPUT -p udp -m udp --dport 123 -j ACCEPT
  iptables -A OUTPUT -p tcp -m tcp --dport 80 -j ext_websites
  iptables -A OUTPUT -j DROP
\end{verbatim}

The OUTPUT chain, allowing :
\begin{itemize}
  \item all local traffic (not leaving the physical PC)
  \item established and related connections
  \item ICMP replies
  \item some remote services (destination IP addresses of hosting servers are not checked)
  \begin{itemize}
    \item dns (udp 53) server
    \item dhcp (udp 67)
    \item smtp (tcp 25)
    \item ntp (udp 123)
  \end{itemize}
  \item external websites (tcp 80), listed in chain ext\_websites
  \item everything else is not allowed (dropped)
\end{itemize}

\subsubsection{icmp in chain}

\begin{verbatim}
  iptables -N icmp_in
  iptables -A icmp_in -p icmp -m icmp --icmp-type 8 -j ACCEPT
  iptables -A icmp_in -p icmp -m icmp --icmp-type 0 -j ACCEPT
  iptables -A icmp_in -p icmp -m icmp --icmp-type 3 -j ACCEPT
  iptables -A icmp_in -p icmp -m icmp --icmp-type 4 -j ACCEPT
  iptables -A icmp_in -p icmp -m icmp --icmp-type 11 -j ACCEPT
  iptables -A icmp_in -p icmp -m icmp --icmp-type 12 -j ACCEPT
  iptables -A icmp_in -j DROP
\end{verbatim}

All allowed incoming ICMP message types.

\subsubsection{admin IP chain}

\begin{verbatim}
  iptables -N admin_IP
  iptables -A admin_IP -s 10.100.2.3 -j ACCEPT
  iptables -A admin_IP -s 10.100.2.4 -j ACCEPT
  iptables -A admin_IP -s 10.100.2.7 -j ACCEPT
  iptables -A admin_IP -j DROP
\end{verbatim}

A list of allowed IP addresses of admin PC's. Everything else is not allowed.

\subsubsection{webmaster IP chain}

\begin{verbatim}
  iptables -N webmaster_IP
  iptables -A webmaster_IP -s 10.100.2.11 -j ACCEPT
  iptables -A webmaster_IP -s 10.100.2.17 -j ACCEPT
  iptables -A webmaster_IP -s 10.100.2.34 -j ACCEPT
  iptables -A webmaster_IP -s 10.100.2.50 -j ACCEPT
  iptables -A webmaster_IP -j admin_IP
\end{verbatim}

A list of allowed IP addresses of webmaster PC's.
At the end of the list, it jumps to the admin\_IP chain, chaining both chains.

\subsubsection{external websites chain}

\begin{verbatim}
  iptables -N ext_websites
  iptables -A ext_websites -d 212.211.132.250 -j ACCEPT
  iptables -A ext_websites -d 212.211.132.32 -j ACCEPT
  iptables -A ext_websites -d 195.20.242.89 -j ACCEPT
  iptables -A ext_websites -d 130.89.149.225 -j ACCEPT
  iptables -A ext_websites -d 86.59.118.153 -j ACCEPT
  iptables -A ext_websites -d 130.89.149.227 -j ACCEPT
  iptables -A ext_websites -d 128.31.0.51 -j ACCEPT
  iptables -A ext_websites -d 86.59.118.153 -j ACCEPT
  iptables -A ext_websites -d 67.228.198.100 -j ACCEPT
  iptables -A ext_websites -d 140.211.166.6 -j ACCEPT
  iptables -A ext_websites -d 140.211.166.21 -j ACCEPT
  iptables -A ext_websites -j LOG
\end{verbatim}

A list of allowed external websites for updates (mirrors of Debian, webmin and Drupal, in this example).
All other requests for external websites are logged. This can be useful for monitoring : notification of abuse, or if you forgot to add an allowed website.i

\section{Conclusion}
